Data_Spatial

### 2020-06-29 src/Data_Spatial.Rmd
I just moved the .Rmd to the .\src and added some spaces to makeit more readable.

#remove below seconds, create columns for date and hour
bb2020$datetext <- strtrim(x = bb2020$datetext, width = 19) 
bb2020$datetext2 <- bb2020$datetext
bb2020$datetext2 <- strtrim(x = bb2020$datetext2, width = 13)
bb2020$datetext2 <- colsplit(bb2020$datetext2, pattern = " ", c("date", "hour"))
bb2020$hour <- bb2020$datetext2$hour
bb2020$date <- bb2020$datetext2$date
#delete datetext2 column(s)
bb2020 <- bb2020[ ,c(1:6,8:9)]
#convert to date form datetext and date columns
bb2020$datetext <- as.POSIXct(strptime(bb2020$datetext, format = "%Y-%m-%d %H:%M:%S", "GMT")) 
bb2020$date <- as.POSIXct(strptime(bb2020$date, format = "%Y-%m-%d", "GMT"))
#add column for month and day of the week to create stats below
bb2020$month <- month(bb2020$date, label = TRUE, abbr = TRUE)
bb2020$weekday <- wday(bb2020$date, label = TRUE)
#create 2 dataframes for 2 matching periods
Wks2019 <- selectByDate(bb2020, start = "2019-01-21", end = "2019-06-02")
names(Wks2019)[1:2] <- c("dateOnly", "date") #413788 obs
Wks2019$Year <- rep(2019)
Wks2020 <- selectByDate(bb2020, start = "2020-01-20", end = "2020-05-31")
names(Wks2020)[1:2] <- c("dateOnly", "date") #521773 obs
Wks2020$Year <- rep(2020)
TSbb19_20 <- rbind(Wks2019, Wks2020) #935561 obs
#exclude obs from Saturdays, Sundays, bank holidays and midnight-6am
TSbb19_20 <- selectByDate(TSbb19_20, day = "weekday", hour = 6:23) #641830 obs
TSbb19_20$dateOnly <- as.character(TSbb19_20$dateOnly)
TSbb19_20 <- TSbb19_20[which(TSbb19_20$dateOnly != "2019-04-19" &
                               TSbb19_20$dateOnly != "2019-04-22" &
                               TSbb19_20$dateOnly != "2019-05-06" &
                               TSbb19_20$dateOnly != "2019-05-27" &
                               TSbb19_20$dateOnly != "2020-04-10" &
                               TSbb19_20$dateOnly != "2020-04-13" &
                               TSbb19_20$dateOnly != "2020-05-08" &
                     TSbb19_20$dateOnly != "2020-05-25"),] #619821 obs
rm(bb2020)
rm(Wks2019)
rm(Wks2020)
```

@@

# get LA 
# (i) directly from the web
#la <- readOGR("http://geoportal1-ons.opendata.arcgis.com/datasets/b6d2e15801de45328b760a4f55d74318_0.geojson?outSR={%22latestWkid%22:3857,%22wkid%22:102100}")#, layer="OGRGeoJSON")
# or, (ii) from the json I saved locally
path.json <- paste(path, "./data/raw/Local_Authority_Districts_(April_2019)_Boundaries_UK_BFE.geojson", sep = "")
la <- readOGR(path.json)#, layer="OGRGeoJSON")
# UK BGC la <- readOGR("https://opendata.arcgis.com/datasets/0e07a8196454415eab18c40a54dfbbef_0.geojson")
# UK BFC la <- readOGR("https://opendata.arcgis.com/datasets/1d78d47c87df4212b79fe2323aae8e08_0.geojson")

# source: https://data.gov.uk/dataset/7c387c64-d25f-474a-b07e-b933578caea2/local-authority-districts-april-2019-boundaries-uk-bfe

@@

```{r include=TRUE, echo=FALSE, results= 'markup', message=FALSE, warning = FALSE, fig.height=600, fig.width=14}
#split by year before create stats by each hour of each day of the week
TS2019 <- selectByDate(TSbb19_20sp, year = 2019) #272248 obs
TS2020 <- selectByDate(TSbb19_20sp, year = 2020) #343268 obs

>>>>

#remove below seconds, create columns for date and hour
bb2020$datetext <- strtrim(x = bb2020$datetext, width = 19) 
bb2020$datetext2 <- bb2020$datetext
bb2020$datetext2 <- strtrim(x = bb2020$datetext2, width = 13)
bb2020$datetext2 <- colsplit(bb2020$datetext2, pattern = " ", c("date", "hour"))
bb2020$hour <- bb2020$datetext2$hour
bb2020$date <- bb2020$datetext2$date

#delete datetext2 column(s)
bb2020 <- bb2020[ ,c(1:6,8:9)]

#convert to date form datetext and date columns
bb2020$datetext <- as.POSIXct(strptime(bb2020$datetext, format = "%Y-%m-%d %H:%M:%S", "GMT")) 
bb2020$date <- as.POSIXct(strptime(bb2020$date, format = "%Y-%m-%d", "GMT"))

#add column for month and day of the week to create stats below
bb2020$month <- month(bb2020$date, label = TRUE, abbr = TRUE)
bb2020$weekday <- wday(bb2020$date, label = TRUE)

#create 2 dataframes for 2 matching periods
Wks2019 <- selectByDate(bb2020, start = "2019-01-21", end = "2019-06-02")
names(Wks2019)[1:2] <- c("dateOnly", "date") #413788 obs
Wks2019$Year <- rep(2019)
Wks2020 <- selectByDate(bb2020, start = "2020-01-20", end = "2020-05-31")
names(Wks2020)[1:2] <- c("dateOnly", "date") #521773 obs
Wks2020$Year <- rep(2020)
TSbb19_20 <- rbind(Wks2019, Wks2020) #935561 obs

#exclude obs from Saturdays, Sundays, bank holidays and midnight-6am
TSbb19_20 <- selectByDate(TSbb19_20, day = "weekday", hour = 6:23) #641830 obs
TSbb19_20$dateOnly <- as.character(TSbb19_20$dateOnly)
TSbb19_20 <- TSbb19_20[which(TSbb19_20$dateOnly != "2019-04-19" &
                               TSbb19_20$dateOnly != "2019-04-22" &
                               TSbb19_20$dateOnly != "2019-05-06" &
                               TSbb19_20$dateOnly != "2019-05-27" &
                               TSbb19_20$dateOnly != "2020-04-10" &
                               TSbb19_20$dateOnly != "2020-04-13" &
                               TSbb19_20$dateOnly != "2020-05-08" &
                     TSbb19_20$dateOnly != "2020-05-25"),] #619821 obs
rm(bb2020)
rm(Wks2019)
rm(Wks2020)
```

@@

# get LA 
# (i) directly from the web
#la <- readOGR("http://geoportal1-ons.opendata.arcgis.com/datasets/b6d2e15801de45328b760a4f55d74318_0.geojson?outSR={%22latestWkid%22:3857,%22wkid%22:102100}")#, layer="OGRGeoJSON")
# or, (ii) from the json I saved locally
path.json <- paste(path, "./data/raw/Local_Authority_Districts_(April_2019)_Boundaries_UK_BFE.geojson", sep = "")
la <- readOGR(path.json)#, layer="OGRGeoJSON")
# source: https://data.gov.uk/dataset/7c387c64-d25f-474a-b07e-b933578caea2/local-authority-districts-april-2019-boundaries-uk-bfe

@@

```{r include=TRUE, echo=FALSE, results= 'markup', message=FALSE, warning = FALSE, fig.height=600, fig.width=14}

#split by year before create stats by each hour of each day of the week
TS2019 <- selectByDate(TSbb19_20sp, year = 2019) #272248 obs
TS2020 <- selectByDate(TSbb19_20sp, year = 2020) #343268 obs
